---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.path = "man/figures/README-",
  out.width = "70%"
)
```

# ednar

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/ednar)](https://CRAN.R-project.org/package=ednar)
<!-- badges: end -->

The goal of ednar is to provide a suite of easy to use functions to help with qPCR 
based targeted eDNA data analysis.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("alexd106/ednar")
```
## Example usage

### Plotting standard curves

To plot a basic qPCR standard curve for a specified target: 

```{r calib_plot1, warning=FALSE, message=FALSE, fig.align='center'}
library(ednar)
calib_plot(calib_data, target = "706")
```

To include the the limit of detection (LOD) and optionally the limit of quantification (LOQ) (usually obtained using the `calib_lod()` function) on the plot use the `lod =` argument. LODs and LOQs can be supplied as either a `data.frame` object or as a vector.  

```{r calib_plot2, warning=FALSE, message=FALSE, fig.align='center'}
# LOD as a vector
lod_vec <- 1.5
calib_plot(calib_data, target = "706", lod = lod_vec)
```


```{r calib_plot3, warning=FALSE, message=FALSE, fig.align='center'}
# LOD and LOQ as a vector
lod_vec <- c(1.5, 4.1)
calib_plot(calib_data, target = "706", lod = lod_vec)
```


```{r calib_plot4, warning=FALSE, message=FALSE, fig.align='center'}
# LOD and LOQ as a tibble
library(tibble)
lod_data <- tibble(Target = "706", lod = 1.5, loq = 4.3)
calib_plot(calib_data, target = "706", lod = lod_data)
```

The `robust = TRUE` argument will exclude standards with less than 50% detections when fitting the linear model

```{r calib_plot5, warning=FALSE, message=FALSE, fig.align='center'}
# LOD and LOQ as a tibble and robust = TRUE
library(tibble)
lod_data <- tibble(Target = "706", lod = 12.5, loq = 16.3)
calib_plot(calib_data, target = "706", lod = lod_data, robust = TRUE)
```

### Summary statistics from standard curve data

To generate a table of summary statistics (slope and intercept estimates, test statistics etc) using standard curve data for each 'Target' we can use the `calib_stats()` function with the argument `type = "effects"`. 

```{r calib_stats1, warning=FALSE, message=FALSE, collapse=TRUE}
sum_stats <- calib_stats(calib_data, type = "effects")
sum_stats
```

To obtain the *R^2* estimates use `type = "r2"`

```{r calib_stats2, warning=FALSE, message=FALSE, collapse=TRUE}
sum_stats <- calib_stats(calib_data, type = "r2")
sum_stats
```

### Estimate LOD and LOQ from standard curve data

To obtain model based estimates of LOD and LOQ from the standard curve data from each 'Target' we can use the `calib_lod()` function. The `calib_lod()` function can automatically determine the 'best' model for LOQ by fitting linear, exponential decay, and up to 6^th order polynomial models and selecting the model with the lowest residual standard error. LOD estimates are obtained by fitting a series of dose response models using the `drc` package (see the `drc::getMeanFunctions()` function to see all possible models) and selecting the 'best' model based on log likelihood values, Akaike's information criterion and residual variance.  

```{r calib_lod1, warning=FALSE, message=FALSE, collapse=TRUE}
qpcr_lod <- calib_lod(data = calib_data, threshold = 0.35,
              lod.fit = "best", loq.fit = "best")
```

To display a summary of the LOD and LOQ estimates for each 'Target' 

```{r calib_lod2, warning=FALSE, message=FALSE, collapse=TRUE}
qpcr_lod$assaySum
```

To display a summary for each of the standards and Targets

```{r calib_lod3, warning=FALSE, message=FALSE, collapse=TRUE}
qpcr_lod$standardsSum
```

With the `calib_lod()` function you can also fit specific models for both LOD and LOQ estimation. For example to estimate LOQ using an exponential decay model and LOD with a 4 parameter Weibull type II model. 
```{r calib_lod4, warning=FALSE, message=FALSE, collapse=TRUE}
qpcr_lod <- calib_lod(data = calib_data, threshold = 0.35,
              lod.fit = "W2.4", loq.fit = "decay")
```
